---
title: "individual2"
output: html_document
---
### Reading in datasets
```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```

```{r}
library(tidyverse)
library(lubridate)
library(ggrepel)
```

```{r}
str(acc)
str(per)
```

## Question 1

### Setting the names from binary
```{r}
acc$DAY_WEEK %>%
  recode("1"="Sunday", "2"="Monday", "3"="Tuesday", "4"="Wednesday", "5"="Thursday", "6"="Friday", "7"="Saturday")
```

```{r}
acc %>%
  group_by(DAY_WEEK) %>%
  summarise(Count=n(), .groups="keep") %>%
  ggplot(aes(x=DAY_WEEK, y=Count)) + geom_bar(stat="identity")
```

After assessing the data, I can conclude that the weekend tends to have more accidents than the weekdays. 


### Question 2

```{r}
fatal <- per %>% 
  filter(INJ_SEV==4) 
```


### Question 3

```{r}
dangerous <- fatal%>%
  drop_na(MAKE) %>%
  group_by(STATE, MAKE) %>%
  summarize(Deaths = sum(PER_NO)) %>%
  group_by(STATE) %>%
  top_n(1, Deaths)
dangerous
```
### Question 4

```{r}
states <- map_data("state")
makeCodes <- read.csv("makeCodes.csv")
stateCodes <- read.csv("stateCodes.csv")
stateCodes$State <- tolower(stateCodes$State)
```

```{r}
dangerous2 <- dangerous %>%
  left_join(stateCodes, by=c("STATE"="Codes")) %>%
  left_join(makeCodes, by=c("MAKE"="Codes")) 
```

```{r}
dangerousStateMap <- dangerous2 %>%
  left_join(states, by=c("State"="region"))
```

```{r}
states2 <- states %>%
  group_by(region) %>%
  summarise(long=mean(long), lat=mean(lat), .groups="keep") %>%
  left_join(dangerous2, by=c("region"="State"))
```
```{r}
ggplot(dangerousStateMap, aes(x= long, y = lat, fill=Deaths)) + geom_polygon(aes(group=group)) +
geom_text_repel(data=states2, color="black", aes(label=Make), size = 1.5) + ggtitle("Map of states with the most dangerous vehicle make")
```

### Question 5

```{r}
perAcc <- per %>%
  left_join(acc, by="ST_CASE") %>%
  distinct(ST_CASE, .keep_all = TRUE)
```

### Question 6

```{r}
perAcc %>%
  ggplot(aes(x=DAY_WEEK, fill =SEX)) +
  geom_bar(position="dodge", stat="count")
  
```


### Question 7

```{r}
library(readxl)
GLCs <- readxl::read_xlsx("GLCs.xlsx")
GLCs <- GLCs %>% 
  select(CountyName, CountyCode, StateName, StateCode)
GLCs$CountyName <- tolower(GLCs$CountyName)
GLCs$StateName <- tolower(GLCs$StateName)
GLCs$CountyCode <- as.numeric(GLCs$CountyCode)
GLCs$StateCode <- as.numeric(GLCs$StateCode)
```



```{r}
county <- map_data("county")


fatal %>%
  group_by(STATE, COUNTY) %>%
  summarise(Deaths = sum(PER_NO)) %>%
  left_join(GLCs, by=c("STATE"="StateCode", "COUNTY"="CountyCode")) %>%
  distinct(COUNTY, STATE, .keep_all = TRUE) %>%
  left_join(county, by=c("CountyName"="subregion", "StateName"="region")) %>%
  ggplot(aes(x= long, y = lat, group=group, fill=Deaths)) + geom_polygon() + ggtitle("Deaths by County")

```



### Question 8

### Winter
```{r}

Winter <- fatal %>%
  filter(MONTH == 12 | MONTH <= 2)
Winter %>%
  select(PER_NO, ST_CASE, STATE) %>%
  group_by(STATE) %>%
  summarise(Deaths=sum(PER_NO))

```

### Summer
```{r}
Summer <- fatal %>%
  filter(MONTH >= 6 & MONTH <= 8)
Summer %>%
  select(PER_NO, ST_CASE, STATE) %>%
  group_by(STATE) %>%
  summarise(Deaths=sum(PER_NO))

```

After comparing the difference between Summer and Winter deaths, I can conclude that Summer has more deaths than the winter.